---
title: "Research_Project"
author: "Brianna MacDonald"
date: "12/10/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Importing packages
library(readr)
library(tidyverse)
library(plotly)
library(ggthemes)
library(gganimate)
library(geosphere)
library(DT)
library(scales)
library(openair)
library(corrplot)
library(viridisLite)
library(viridis)
library(RColorBrewer)
library(ggdark)
library(anytime)
library(usmap)
library(covid19nytimes)
library(sigr)
library(caret)
library(glmnet)
library(Metrics)
```

## Initial Visualization of Worldwide Number of Cases

```{r cases_worldwide}
# Importing data
ts_confirmed <- read_csv(file = "time_series_covid_19_confirmed.csv",
                         col_types = cols(
                           .default = col_double(),
                           `Province/State` = col_character(),
                           `Country/Region` = col_character()
                         ))

ts_recovered <- read_csv(file = "time_series_covid_19_recovered.csv",
                         col_types = cols(
                           .default = col_double(),
                           `Province/State` = col_character(),
                           `Country/Region` = col_character()
                         ))

ts_deaths <- read_csv(file = "time_series_covid_19_deaths.csv",
                      col_types = cols(
                        .default = col_double(),
                        `Province/State` = col_character(),
                        `Country/Region` = col_character()
                      ))

codes <- read_csv('https://raw.githubusercontent.com/plotly/datasets/master/2014_world_gdp_with_codes.csv',
                  col_types = cols(
                    COUNTRY = col_character(),
                    `GDP (BILLIONS)` = col_double(),
                    CODE = col_character()
                  ))
# Preprocessing Data
ts_confirmed <- ts_confirmed %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

  gather("Date", "Confirmed", -c("Province/State", "Country/Region", "Lat", "Long")) %>%
ts_recovered <- ts_recovered %>%
  gather("Date", "Recovered", -c("Province/State", "Country/Region", "Lat", "Long")) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

ts_deaths <- ts_deaths %>%
  gather("Date", "Deaths", -c("Province/State", "Country/Region", "Lat", "Long")) %>%
  mutate(Date = as.Date(Date, "%m/%d/%y"))

# Organizing data
ts_total <- ts_confirmed %>%
  left_join(ts_deaths) %>%
  left_join(ts_recovered) %>%
  mutate(Recovered = replace_na(Recovered, replace = 0))

## We all know "Diamond Princess" and "MS Zaandam" are cruises, So we have to remove them from the data

ts_total <- ts_total %>%
  filter(`Country/Region` != "Diamond Princess") %>%
  filter(`Country/Region` != "MS Zaandam")

ts_total$Deaths[is.na(ts_total$Deaths)] <- 0

## Created a dataset including latest news of COVID-19

cases_latest <- ts_total %>%
  group_by(`Country/Region`, Date) %>%
  summarise(Confirmed  = sum(Confirmed),
            Recovered = sum(Recovered),
            Deaths = sum(Deaths)) %>%
  mutate("New Cases" = Confirmed - lag(Confirmed, 1) ) %>%
  filter(Date == max(Date))

day_latest <- max(cases_latest$Date)

cases_total_date <- ts_total %>%
  rename(Region = `Country/Region`) %>%
  group_by(Date) %>%
  summarise(Confirmed = sum(Confirmed),
            Deaths = sum(Deaths),
            Recovered = sum(Recovered)) %>%
  mutate("New_Cases" = Confirmed - lag(Confirmed, 1))

cases_total_date$New_Cases[is.na(cases_total_date$New_Cases)] <- 0 

cases_total_latest <- cases_total_date %>%
  filter(Date == max(Date))

# Visualization
cases_all <- cases_total_date %>%
  select(-Confirmed, -New_Cases) %>%
  gather("Status", "Cases", -"Date")

barchart <- ggplot(data = cases_total_date, aes(x = Date)) +
  geom_bar(aes(y = Confirmed), position = "stack", stat = "identity", fill = "#ff5050") +
  geom_bar(data = cases_all, aes(y = Cases, fill = Status), position = "stack", stat = "identity") +
  scale_fill_manual(values = c("#000000", "#009900")) +
  scale_y_continuous(breaks = seq(0, 21000000, by = 1000000), labels = comma) +
  theme_solarized(base_size = 10, light = TRUE)+
  theme(plot.margin = margin(0, 0, 0, 0, "pt"),
        panel.background = element_rect(fill = "White"),
        legend.position = "bottom",
        axis.title = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank()) +
  ggtitle("World COVID-19 Total Cases by Day")

ggplotly(barchart) %>%
  layout(legend = list(orientation = 'h'))
```

## Correlations

```{r case_corr}
# Mortality rate and recovery rate
cases_total_date <- cases_total_date %>%
  group_by(Date, Confirmed) %>%
  mutate(Mortality_rate = Deaths / Confirmed,
         Recovery_rate = Recovered / Confirmed) %>%
  ungroup()

# Correlations
cases_total_date %>%
  select(-Date) %>%
  na.omit() %>%
  cor(use = "pairwise.complete.obs") %>%
  corrplot.mixed(tl.col = "black", tl.pos = "d", tl.cex = 0.7, cl.cex = 0.7,
                  number.cex = 0.7)

```

## US Cases by State, March
```{r us_cases_by_state_march}
# March 15th, August 15th, November 15th
us <- dt %>%
    filter(Country == "US") %>%
    mutate(Province = case_when(
           Province == "Chicago"~"Illinois",
           TRUE~Province)) %>%
    filter(Province!= "Recovered")

pop <- read_csv("US_State_Population.csv",
        col_types = cols(Province = col_character(), pop2020 = col_number(), lat = col_number(), lon = col_number())) %>%
        mutate(Province = case_when(
                Province == "U.S. Virgin Islands"~"Virgin Islands",
                TRUE~Province))
# March 15th
us_map <- us %>% left_join(pop, by = "Province") %>%
    filter(Date == anydate("03/15/2020"))
# Mapping
us_map <- us_map %>%
    mutate(confirmed_per_capita = Confirmed/pop2019)

teal <- "#008080" #RGB color code

plot_usmap(data = us_map, values = "Confirmed", color = teal, labels = TRUE) +
    scale_fill_continuous(low = "white", high = teal, name = "Confirmed Cases", label = scales::comma) +
    theme(legend.position = "right") +
    theme(panel.background = element_rect(colour = "black")) +
    labs(title = "Confirmed", caption = "Confirmed Cases as of 15 March ")

plot_usmap(data = us_map, values = "confirmed_per_capita", color = teal, labels = TRUE) +
    scale_fill_continuous(low = "white", high = teal, name = "Confirmed Cases", label = scales::comma) +
    theme(legend.position = "right") +
    theme(panel.background = element_rect(colour = "black")) +
    labs(title = "Confirmed per Capita", caption = "Confirmed Cases per Capita as of March 15th ")
```
## US Cases by State, August
```{r us_cases_by_state_august}
# August 15th
us_map <- us %>% left_join(pop, by = "Province") %>%
    filter(Date == anydate("08/15/2020"))
# Mapping
us_map <- us_map %>%
    mutate(confirmed_per_capita = Confirmed/pop2019)

plot_usmap(data = us_map, values = "Confirmed", color = teal, labels = TRUE) +
    scale_fill_continuous(low = "white", high = teal, name = "Confirmed Cases", label = scales::comma) +
    theme(legend.position = "right") +
    theme(panel.background = element_rect(colour = "black")) +
    labs(title = "Confirmed", caption = "Confirmed Cases as of 15 August ")

plot_usmap(data = us_map, values = "confirmed_per_capita", color = teal, labels = TRUE) +
    scale_fill_continuous(low = "white", high = teal, name = "Confirmed Cases", label = scales::comma) +
    theme(legend.position = "right") +
    theme(panel.background = element_rect(colour = "black")) +
    labs(title = "Confirmed per Capita", caption = "Confirmed Cases per Capita as of August 15th ")
```
## US Cases by State, November
```{r us_cases_by_state_nov}
# November 15th
us_map <- us %>% left_join(pop, by = "Province") %>%
    filter(Date == anydate("11/15/2020"))
# Mapping
us_map <- us_map %>%
    mutate(confirmed_per_capita = Confirmed/pop2019)

plot_usmap(data = us_map, values = "Confirmed", color = teal, labels = TRUE) +
    scale_fill_continuous(low = "white", high = teal, name = "Confirmed Cases", label = scales::comma) +
    theme(legend.position = "right") +
    theme(panel.background = element_rect(colour = "black")) +
    labs(title = "Confirmed", caption = "Confirmed Cases as of 15 November ")

plot_usmap(data = us_map, values = "confirmed_per_capita", color = teal, labels = TRUE) +
    scale_fill_continuous(low = "white", high = teal, name = "Confirmed Cases", label = scales::comma) +
    theme(legend.position = "right") +
    theme(panel.background = element_rect(colour = "black")) +
    labs(title = "Confirmed per Capita", caption = "Confirmed Cases per Capita as of November 15th ")
```
## How has quarantine affected the spred of COVID-19?
Looking at the following states of interest: New York, Washington, Rhode Island, Florida, California for number of confirmed deaths before and after quarantine.
```{r quar_before_after_deaths}
# Refreshing data
covid19nytimes_states <- refresh_covid19nytimes_states()

covid19nytimes_counties <- refresh_covid19nytimes_counties()
# Viewing data by states
covid19nytimes_states %>%
  filter(location %in% c("New York", "Illinois", "California", "Texas")) %>%
  filter(data_type == "deaths_total") %>%
  ggplot(aes(x = date, y = value, color = location)) +
  geom_line() +
  theme_minimal(base_size=14) +
  scale_y_continuous() +
  ggtitle("Confirmed Deaths in US States")
```
Looking at the following states of interest: New York, Washington, Rhode Island, Florida, California for number of confirmed cases before and after quarantine.
```{r quar_before_after_cases}
# Refreshing data
covid19nytimes_states <- refresh_covid19nytimes_states()

covid19nytimes_counties <- refresh_covid19nytimes_counties()
# Viewing data by states
covid19nytimes_states %>%
  filter(location %in% c("New York", "Illinois", "California", "Texas")) %>%
  filter(data_type == "cases_total") %>%
  ggplot(aes(x = date, y = value, color = location)) +
  geom_line() +
  theme_minimal(base_size=14) +
  scale_y_continuous() +
  ggtitle("Confirmed Cases in US States")
```
## Regression Model: Chicago
```{r predicting_covid1}
set.seed(143)
chicago <- read.csv("COVID-19_Daily_Testing.csv")
head(chicago, n=10)
chicago$Tests <- gsub(',', '', chicago$Tests)
chicago$Cases <- gsub(',', '', chicago$Cases)
chicago$Tests <- as.numeric(chicago$Tests)
chicago$Cases <- as.numeric(chicago$Cases)
print(chicago$Cases)
 
# Splitting the data into test and train, 80% split/20
row.number <- sample(1:nrow(chicago), 0.8*nrow(chicago))
train = chicago[row.number,]
test = chicago[-row.number,]
dim(train)
dim(test)
# Regression
lmchicago <- lm(chicago$Cases~chicago$Tests, data = train)
wrapFTest(lmchicago)
```
## Predicting values for our Regression Model
```{r pred_lr}
prediction <- predict(lmchicago)
ggplot(lmchicago, aes(x = prediction, y = chicago$Cases)) + 
  geom_point() +
  geom_abline(color = "blue")
```
## Logistic Linear Regression
```{r log}
glm.fit <- glm(chicago$Cases~chicago$Tests, data = train, family = binomial)
summary(glm.fit)
glm.probs <- predict(glm.fit)
glm.probs[1:5]
```
## Polynomial Regression
```{r poly}
plot(chicago$Cases, chicago$Tests)

model_1 <- lm(chicago$Cases ~ poly(chicago$Tests, 1))

# Summary statistics for model_1
summary(model_1)

model_2 <- lm(chicago$Cases ~ poly(chicago$Tests, 2))  

# Summary statistics for model_2
summary(model_2)

model_3 <- lm(chicago$Cases ~ poly(chicago$Tests, 3))

# Summary statistics for model_3
summary(model_3)

model_4 <- lm(chicago$Cases ~ poly(chicago$Tests, 4))

# Summary statistics for model_3, R2 of 87.01, 2.2e-16 p-value
summary(model_4)
# Predictions for this model
poly_pred <- predict(model_4)
```
## Examining best model *model_4*
```{r model_4}
ggplot(model_4, aes(x = poly_pred, y = chicago$Cases)) + 
  geom_point() +
  geom_abline(color = "red")
```
## Evaluating Model
```{r eval}
# First model
RMSE(chicago$Cases, prediction)
R2(chicago$Cases, prediction)
mae(chicago$Cases, prediction)
# Second model
RMSE(chicago$Cases, poly_pred)
R2(chicago$Cases, poly_pred)
mae(chicago$Cases, poly_pred)
wrapFTest(model_4)
```
## Comparing models
```{r compare}
anova(lmchicago, model_4)
```
## Exploring more about Illinois
```{r illinois}
library(ggplot2)
illinois <- read.csv("latest_IL.csv")
head(illinois, n=10)
total_cases <- illinois$total_cases
total_deaths <- illinois$total_deaths
dates <- illinois$date
ggplot(aes(x = dates, y = total_cases), data = illinois) +
  geom_line() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  scale_y_continuous() +
  ggtitle("Confirmed Deaths in Illinois")
```
## Linear Regresssion for Future Illinois Cases
```{r illinois_reg}
set.seed(153)

# Regression
lmillinois <- lm(total_cases~total_deaths, data = illinois)
wrapFTest(lmillinois)
```
## Predicting values for our Regression Model
```{r pred_lr}
prediction_il <- predict(lmillinois)
ggplot(lmillinois, aes(x = prediction_il, y = total_cases)) + 
  geom_point() +
  geom_abline(color = "green")
```
## Evaluating Model
```{r eval}
# Evaluating model
RMSE(total_cases, prediction_il)
R2(total_cases, prediction_il)
mae(total_cases, prediction_il)
anova(lmillinois)
```
## Polynomial Regression Illinois
```{r poly_il}
model_5 <- lm(total_cases ~ poly(total_deaths, 4))
summary(model_5)
# Predictions for this model
poly_pred_il <- predict(model_5)
```
## Plotting poly illinois
```{r plot_poly_il}
ggplot(model_5, aes(x = poly_pred_il, y = total_cases)) + 
  geom_point() +
  geom_abline(color = "green")
```
## Evaluating poly illinois model
```{r eval}
# Evaluating model
# Linear model
RMSE(total_cases, prediction_il)
R2(total_cases, prediction_il)
mae(total_cases, prediction_il)
# Polynomial Model
RMSE(total_cases, poly_pred_il)
R2(total_cases, poly_pred_il)
mae(total_cases, poly_pred_il)
```
## Sample Data
```{r sampling_data}
illinois <- read.csv("latest_IL.csv")
summary(illinois, n=10)
```
## Training/Predicting with Regression Trees
```{r regr_trees}
# Splitting the data into test and train, 80% split/20
row.number <- sample(1:nrow(illinois), 0.8*nrow(illinois))
train = illinois[row.number,]
test = illinois[-row.number,]
summary(train)
```
## Building model for predicting non-positive cases
```{r building_model_cases}
num_folds <- trainControl(method = "cv", number = 5) # Specify 5-fold cross-validation.
parameter_grid <- expand.grid(.cp = 0)
# Predicting Confirmed Cases
il_1 <- read.csv("COVID-19_Daily_Testing.csv")
il_1$People.Not.Positive...Total <- as.numeric(il_1$People.Not.Positive...Total)
il_1$Tests <- as.numeric(il_1$Tests)
row.number <- sample(1:nrow(il_1), 0.8*nrow(il_1))
train_il1 = il_1[row.number,]
test_il1 = il_1[-row.number,]
summary(train_il1)
# Making model
cases_model <- train(People.Not.Positive...Total ~ Tests,  
  data = train_il1, 
  method = "rpart", # CART algorithm
  trControl = num_folds, 
  tuneGrid = parameter_grid,
  na.action=na.exclude
)

print(cases_model)
```
## Building model for predicting fatalities
```{r building_model_deaths}
# Predicting new deaths
il_2 <- read.csv("COVID-19_Daily_Cases__Deaths__and_Hospitalizations.csv")
row.number <- sample(1:nrow(il_2), 0.65*nrow(il_2))
train_il = il_2[row.number,]
test_il = il_2[-row.number,]
# Making model
death_model <- train(
  Deaths...Total ~ Cases...Total + Hospitalizations...Total,  
  data = train, 
  method = "rpart", # CART algorithm
  trControl = num_folds, 
  tuneGrid = parameter_grid,
  na.action=na.exclude
)

print(death_model)
```
## Predicting Positive Cases in Females Aged 40-49 using Regression Trees
```{r 18_24}
# Splitting the data into test and train, 65% split/35
summary(chicago)
split <- initial_split(chicago, prop = .65)
train = chicago[split,]
test = chicago[-split,]
summary(train)
# Predicting new deaths
cases_18 <- train(
  Cases ~ People.Tested...Age.40.49 + 
  People.Tested...Female,  
  data = train, 
  method = "rpart", # CART algorithm
  trControl = num_folds, 
  tuneGrid = parameter_grid
)
print(cases_18)

```
```{r delete}
summary(cases_18)
```
## Plotting Positive Cases in Females Aged 40-49 using RPart
```{r cases_rpart}
library(rpart.plot)
rpart_18 <- rpart(
  People.Positive...Age.40.49 ~ People.Tested...Age.40.49 + 
  People.Tested...Female,  
  data = train, 
  method  = "anova"
  )
# Predictions for this model
rpart.plot(rpart_18)
```